---
title: "Sampling"
---

Follow along with the demo to learn how to sample data from a population using R and the Tidyverse. You should get started by loading the tidyverse in a new R block.

### Task 1: Load the tidyverse

```{r}
library(tidyverse)
```

### Task 2: Load 1,000 rows of data with the ages of people from 18-80

```{r}
set.seed(5678)

population <- tibble(
  age= sample(18:80, 1000, replace = TRUE)
)
```
##replace mean to run it from the start
##sample creates random sample of ages 18-80

### Task 3: Sample 100 rows of data from the dataset

```{r}
sample100 <- population |>
  sample_n(100)
```

##sample_n creates a sample of your population 

### Task 4: Calculate the mean age and standard deviation of the sample

```{r}
sample100 |>
  summarize(
    mean_age = mean(age),
    sd_age = sd(age)
  )

```

### Task 5: Calculate the mean age and standard deviation of the population

```{r}
population |>
  summarize(
    mean_age = mean(age),
    sd_age = sd(age)
  )
  

```

### Task 6: Compare the sample and population statistics

Compare the means and standard deviations. Are they similar? Go back and re-run the sampling and statistics calculations with a smaller sample size. How do the results change?

##The means and sd's for sample and population are pretty similar. They differ by less than 2 values. For a sample size of 50, my means stay similar but the sd change by a few values. For a sample size of 20, my sd stays similar and the mean changes by a few values. 

When finished, save your work, switch to GitHub Desktop, then add, commit and push your changes to GitHub and submit the URL of the notebook in ELMS.
