---
title: "HW2 Prince George's County Crime"
---

In this assignment, we will be working with a dataset reporting the number of crimes by month in Prince George's County, Maryland, for every month in 2024. 

### Task 1: Load the Tidyverse

You will need to load the tidyverse library

```{r}
library(tidyverse)
```

Then you'll load the data from the web. The URL for the data is: https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/pg_crime_totals_2024.csv. You will create a dataframe called `pg_crime` that contains the data from this URL.

### Task 2: Load the Data and Describe It

```{r}
pg_crime <- read_csv("https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/pg_crime_totals_2024.csv")

```
### Task 3: Describe the Data

You should examine that dataframe carefully by clicking on it in the Environment tab in RStudio. Write a couple of sentences describing the categories of incidents. From a reporting standpoint, are any of the categories more important than others? Why or why not? Do any of them make the job of summarizing the data more difficult? Why or why not?

## The categories of incidents include crime totals in Prince George's County from 2024. The incident types are organized in alphabetical order. The incidents with the highest numbers all relate to cars. Accidents, Auto Stolen and Theft from Auto have higher numbers than most other incidents reporter. These categories are important if you are reporting about Auto Theft issues in PG County. From what I can see, I don't see any of them making it hard to summarize data. 

### Task 4: Calculating Averages

Add up the total number of incidents per month for each category using `mutate()`, calling the new column `total_incidents`. Then calculate the average number of incidents per month for each category of crime using the total number of months, calling the new column `average_incidents`. Assign the result to a new dataframe called `pg_crime_averages`.

## Follow this pattern given by Prof. Willis: new_dataframe <- old_dataframe |> mutate(new_column = some formula)

```{r}
total_incidents <- pg_crime |> mutate(total_incidents = `Jan` + `Feb` + `Mar` + `Apr` + `May` + `Jun` + `Jul` + `Aug` + `Sep` + `Oct` + `Nov` + `Dec`)


pg_crime_averages <- total_incidents|> mutate(average_incidents = total_incidents / 12)

```

### Task 5: Describe the Averages

Examine the new dataframe, `pg_crime_averages`, by clicking on it in the Environment tab in RStudio. Write a couple of sentences describing the average number of incidents per month for each category. Which categories have the highest average number of incidents per month? Which have the lowest? What does this tell you about the pattern of crime in Prince George's County? Are there categories that you have questions about? Why or why not?

## On average for this dataset, "Auto Accidents" have the highest average numner of incidents per month. This is also shared by "Theft from Auto" and "Auto Stolen." The incidents with the lowest are "Robbery Residential" (N/A classification), "Sex Offense," and "Homicide." This data tells us that auto related crime incidents are the most prevalent crime incidents reported in Maryland during 2024. I'm curious why "Robbery Residential" is N/A. I also want to know why there are so many auto related crime incidents reported during this time period. 

### Task 6: Testing a Theory

One theory about crime is that it is more likely to occur in the summer months. Test this theory by creating three new columns using `mutate()` that calculate the percentage change compared to the average for each category of crime for the months of June, July, and August. Call the new columns `june_change`, `july_change`, and `august_change`. Assign the result to a new dataframe called `summer_crime_averages`.

## Follow this pattern given by Prof. Willis: new_dataframe <- old_dataframe |> mutate(new_column = some formula)

```{r}
summer_crime_averages <- pg_crime_averages |> mutate(june_change = (Jun- average_incidents)/Jun * 100, july_change = (Jul- average_incidents)/Jul * 100, august_change = (Aug- average_incidents)/Aug * 100)

```

### Task 7: Describe the Summer Crime Averages

Look at the `summer_crime_averages` dataframe by clicking on it in the Environment tab in RStudio. Write a couple of sentences describing the percentage change in the average number of incidents per month for each category of crime for the months of June, July, and August. Do you see any patterns? What does this tell you about the theory that crime is more likely to occur in the summer months?

##Through the summer month percentage change data, I saw that assault shootings had over a 25% change from the average in 2024. In addition, "Auto Stolen" had a positive percentage change from the average, but the data did not go above over 10% for any of the three months. This tells me that Auto Stolen did go up, but did not have a drastic change in the summer months. I also noticed that "B&E Other" had a negative percentage change during the summer months. This tells me that B&E other is less likely during the summer months. Based on all of the data (including the Homicide changes), I am not convinced that the theory that crime is more likely to occur in the summer months is true. 

### Task 8: What's the Story?

Write up a lede and nutgraf of a story focused on the crime data in Prince George's County. It could be from any of the dataframes you've created in this assignment. The lede should be a single sentence that summarizes the most important information in the data. The nutgraf should be a paragraph that provides more context and explains why the data is important.

## LEDE: The frequency of auto related crimes in Prince George's County outweigh other reported crime incidents, according to a report documenting crime data from 2024. 

## NUTGRAF: In Prince George's County, car accidents take the top spot for crime incidents with 5,576 total accidents reported in 2024. Behind accidents, other auto related crime incidents with the highest number of reports include stolen automobiles with 4,790 reported incidents, as well as theft from auto with 4,703 reported incidents. This data collected by the county may support resident concerns over the high prevalence of auto related crimes in the area. 

### Task 9: Save Your Work and Put it on GitHub

Save this notebook, then commit it to your local repository. Push the changes to your GitHub repository and submit the URL for this file in ELMS.
