---
title: "HW3 WH Salaries"
---

In this assignment, we will be working with a dataset containing information about the salaries of White House staffers as of July 2026.

### Task 1: Load the Tidyverse

You will need to load the tidyverse library

```{r}
library(tidyverse)
```

Then you'll load the data from the web. The URL for the data is: https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/white_house_salaries_2026.csv. You will create a dataframe called `wh_salaries` that contains the data from this URL.

### Task 2: Load the Data and Describe It

```{r}
wh_salaries <- read_csv("https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/white_house_salaries_2026.csv")
```
### Task 3: Describe the Data

You should examine that dataframe carefully by clicking on it in the Environment tab in RStudio. Write a couple of sentences describing the categories of incidents. From a reporting standpoint, what stands out to you about this data? Are there any salary values that will cause problems for analysis?

ANSWER HERE
##The data is categorized by name of employee, employee status, employee salary, pay basis and position title. The statuses mentioned are employee, detailee and part-time detailee. All of the employees listed in the data seem to be paid annually. Some employees are listed as having salaries of $0 per year. These salary values could cause problems for analysis because it could skew certain data calculations, such as the average. 

### Task 4: Calculate the Mean and Median of the Salaries

Write code to calculate the mean and median of the salaries in the dataset using the `summarize()` function. You don't need to make a new dataframe, just display the results.

```{r}
wh_salaries |> summarize(mean_salary = mean(SALARY), median_salary = median(SALARY))

```

### Task 5: Add in the minimum and maximum salaries

Take the code you just wrote and add in the minimum and maximum salaries (the functions for that are `min()` and `max()`). Your results should show four columns: mean_salary, median_salary, min_salary, max_salary.

```{r}
wh_salaries |> summarize(mean_salary = mean(SALARY), median_salary = median(SALARY), min_salary = min(SALARY), max_salary = max(SALARY))

```

### Task 6: Describe the Results

Write a couple of sentences describing the results of your analysis. What do the mean and median salaries tell you about the data? What do the minimum and maximum salaries tell you? Are there any outliers in the data?

##The mean salary is $112702.7 and the median salary is $109243. The mean averages all of the salaries reported, while the median is the middle salary value of all salaries reported. The min salary is $0 and the max is $197200. The min and max salaries tell me that there is a very broad range of salaries reported. Based on the mean and median, the min salary of $0 seems to be an outlier because it deviates very far from the mean salary of $112702.7. My best guess is that the employees paid $0 per year are paid through some other government agency. 

### Task 7: Story Idea

Based on the data, what is your best story idea? What is the next step you would take to develop that story idea?

##The employees who make $0 a year intrigue me. As I mentioned earlier, I assume they are employed/ paid by some other agency. My question would be why they are stationed at the white house. I am also curious as to why there is a dedicated Faith office in the WH. The next step would be looking into the employees themselves and seeing where they might be paid from (other agency payroll information).
