---
title: "HW4 MD CityCrime"
---

In this assignment, we will be working with a dataset of total crimes reported by Maryland cities for two consecutive years. You'll load the data, calculate the violent and property crime rates for each city for both years, and then calculate the percentage change in the crime rates between the two years. You'll also write a short story based on your findings.

### Task 1: Load the Tidyverse and Janitor

You will need to load the tidyverse library and the janitor library

```{r}
library(tidyverse)
```

Then you'll load the data from the web. The URL for the data is: https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/md_cities_crime_2019_2020.csv. You will create a dataframe called `md_cities` that contains the data from this URL.

### Task 2: Load the Data

```{r}
crimes <- read_csv("https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/md_cities_crime_2019_2020.csv")
```
### Task 3: Calculating Rates

Calculate violent and property crime rates for each city for both years using `mutate()`. You'll create four new columns: `violent_rate_2019`, `property_rate_2019`, `violent_rate_2020`, and `property_rate_2020`. You'll need to figure out the proper unit rate, given that there are varying populations. There's no perfect choice; choose the one that works best for you. Save the results to a new dataframe called `md_cities_rates`.

```{r}
md_cities_rates <- crimes |> mutate(
violent_rate_2019 = (violent2019/pop2019) * 10000 ,
property_rate_2019 = (property2019 /pop2019) * 10000 ,
violent_rate_2020 = (violent2020 /pop2020) * 10000  ,
property_rate_2020 = (property2020 / pop2020) * 10000) 

```

### Task 4: Describe the Rates

Examine the new dataframe, `md_cities_rates`, by clicking on it in the Environment tab in RStudio and sorting by the rate columns. Write a couple of sentences about the results, comparing and contrasting the cities with the highest rates (both violent and property) in 2019 and 2020. Do you see any patterns? What does this tell you about crime in Maryland cities? What's the story here?

ANSWER HERE
Ocean City had the highest rates for Property rate 2019, violent rate 2020 and property rate 2020. Baltimore had the highest rate for violent rate 2019. Baltimore City had the highest rates for all 4 colums in the normal crimes dataset. This tells us that Baltimore City has the highest crime rates but Ocean City had the highest for three categories per population of 10000 people. The story could be that Baltimore City and Ocean City have the highest crime rates based on the population. 


### Task 5: Calculate Percentage Change

Let's add two percentage change columns to your dataframe: one for violent crime rates and one for property crime rates. You'll calculate the percentage change between 2019 and 2020 for each city. Save the results to a new dataframe called `md_cities_change`.

```{r}
md_cities_change <- md_cities_rates |> mutate(pct_change_property = (`property_rate_2020`- `property_rate_2019`)/`property_rate_2019` * 100 , pct_change_violent= (`violent_rate_2020`- `violent_rate_2019`)/`violent_rate_2019` * 100 )

```

##I'm not sure I did this one correctly. 

### Task 7: Describe Percentage Changes

Look at the `md_cities_change` dataframe by clicking on it in the Environment tab in RStudio. Write a couple of sentences describing the percentage change results. What's more interesting here, the patterns or the outliers? What's the story in the data?

ANSWER HERE
Aberdeen had a decrease in property crime rates from 2019 to 2020 around 8% and violent crime rates from 2019 to 2020 around 11%. Some of the cities like Cecilton had a 100% decrease in violent crimes from 2019 to 2020. Charlestown came back as "inf."

### Task 8: Tell the Best Story

Write up a lede and nutgraf of a story with the _best_ story idea you've found in this assignment. The lede should be a single sentence that summarizes the most important information in the data. The nutgraf should be a paragraph that provides more context and explains why the data is important.

##Lede: Ocean City, Md. has the highest property crime rates per 10,000 in 2019 and 2020. 

##Nutgraf: During 2019 and 2020, Maryland crime rates revealed that Ocean City, Md. was found to have the highest property crime rates per 10,000 people. In 2019, Ocean City had around 1036 property crimes per 10,000 people. In 2020, Ocean City had around 1212 property crimes per 10,000 people. For both years, Ocean City outranked Baltimore city in property crimes per 10,000 people. 

### Task 9: Save Your Work and Put it on GitHub

Save this notebook, then commit it to your local repository. Push the changes to your GitHub repository and submit the URL for this file in ELMS.


##Task 10

Using the 2020 violent crime rate, calculate the percentiles using `ntile()`. Save the results to a new dataframe called `md_cities_percentiles`. How many cities are above the 90th percentile?

##11 cities are above the 90th percentile.

```{r}
md_cities_percentiles <- md_cities_change |> mutate(percentile = ntile(violent_rate_2020, 100))
```

Calculate the Interquartile Range of the 2020 violent crime rate using `summarize()` and R's `IQR()` function. The syntax is similar to how `mean()` works.

```{r}
md_cities_change |> summarize(IQR(violent_rate_2020))
```
Do the results change your best story idea from the previous assignment? Why or why not?
## I don't think that this IQR result changes my story idea from the previous assigment. Instead, I think it is a good comparison to show that the crime rate results for Ocean City are outliers when you factor in crimes per 10,000 people. 

